Fast and adaptive cointegration based model for forecasting high frequency financial time series

### 2017-01-16 articles-conferences/iscef2016/body.tex
specific comments 4,5

Cointegration vectors can be found applying the Johansen method which uses a
sample of the last historical data. However, VECM assumes cointegration vectors
do not change in time.  In fact, \cite{gregoryETal1996} addresses that the
long-run relationships between the time series  might change due to several
economic factors that can lead to structural breaks in the cointegration
relationship.  In order to show that the number of cointegration vectors depends
on the amount $L$ of historical data and the number of lags $p$ in the VECM, we
used a grid search.  We defined a grid of possible values for $L$ and $p$. $L$
goes throughout $[2,14]$ hours (1 $hour = 360$ data points) with a step size of
4 hours and $p$ throughout $[1,5]$ with step size of 1. The first experiment
consisted in determining the number of cointegration vectors for all
combinations of $L$ and $p$.  We used four forex rates: EURUSD, GBPUSD, USDCHF
and USDJPY with 10-second frequency. Data started at 13:00 GMT of the 13th of
August 2014, when the New York and London financial markets opened.

Figure \ref{fig:hists} shows the distribution of the number of cointegration
vectors given by the Johansen method for different values of $L=[2,6,10,14]$
hours  and $p=1$ . This procedure was carried out by a sliding window of
historical data moving 1000 times.

@@

>>>>

Cointegration vectors can be found applying the Johansen method which uses a
sample of the last historical data. However, VECM assumes cointegration vectors
do not change in time.  In fact, \cite{gregoryETal1996} addresses that the
long-run relationships between the time series  might change due to several
economic factors that can lead to structural breaks in the cointegration
relationship.  In order to show that the number of cointegration vectors depends
on the amount $L$ of historical data and the number of lags $p$ in the VECM, we
used a grid search.  We defined arbitrarily a grid of possible values for $L$
and $p$. $L$ goes throughout $[2,14]$ hours (1 $hour = 360$ data points) with a
step size of 4 hours and $p$ throughout $[1,5]$ with step size of 1. The idea is
to show the variability of the number of cointegration vectors when we changed
these two parameters. We used four forex rates: EURUSD, GBPUSD, USDCHF and
USDJPY with 10-second frequency. Data started at 13:00 GMT of the 13th of August
2014, when the New York and London financial markets opened.

@@

Figure \ref{fig:hists} shows the distribution of the number of cointegration
vectors given by the Johansen method for different values of $L=[2,6,10,14]$
hours and $p=1$ . This procedure was carried out by a sliding window of
historical data moving 1000 times. We can observe that the distribution of
cointegration vectors is always changing with different values of $L$. When
$L=2$ hours, there were no cointegration in more than 60\% of the iterations.
Cointegration increases when $L=6$ hours and is maximum when $L=10$ hours where
one cointegration vectors was found for all 1000 iterations. The ocurrence of
cointegration starts to decrease with $L=14$ hours.
